---
title: "UpdatedCensusHarmonised"
author: "Cameron Ward"
date: '2023-06-27'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
require(data.table)
require(tidyverse)
choose.files()
LSOA2011 <- read.csv("C:\\Users\\sgabalog\\Documents\\P1\\LSOA_(2011)_to_LSOA_(2021)_to_Local_Authority_District_(2022)_Exact_Fit_Lookup_for_EW_(V3).csv", fileEncoding="UTF-8-BOM") %>%
  select(1:7) %>%
  rename(CHG21 = CHGIND)

df <- LSOA2011

df <- df %>%
  mutate(LSOAXXCD = NA) %>%
  mutate(LSOAXXCD = ifelse(CHG21 == "U", LSOA21CD, LSOAXXCD)) %>%
  mutate(LSOAXXCD = ifelse(CHG21 == "M", LSOA21CD, LSOAXXCD)) %>%
  mutate(LSOAXXCD = ifelse(CHG21 == "S", LSOA11CD, LSOAXXCD))
```


```{r}
df |>
  filter(df$CHG21 == 'U')
```


```{r}
df |>
  filter(df$CHG21 == 'M')
```

```{r}
df |>
  filter(df$CHG21 == 'S')
```

```{r}
df |>
  filter(is.na(LSOAXXCD))
```


```{r}
df |>
  filter(is.na(LSOAXXCD)) |>
  select(CHG21) |>
  group_by(CHG21) |>
  summarise(n = n()) |>
  print(n = Inf)
```




```{r}
length(which(is.na(df$LSOAXXCD)))
```

```{r}
install.packages("RANN")
require(RANN)
choose.files()

pts11 <- read.csv("C:\\Users\\sgabalog\\Documents\\P1\\LSOA_Dec_2011_PWC_in_England_and_Wales_2022_1923591000694358693.csv") %>%
  select(x, y, LSOA11CD) %>%
  semi_join(df %>% filter(is.na(LSOAXXCD)), by = "LSOA11CD")

pts21 <- read.csv("C:\\Users\\sgabalog\\Documents\\P1\\LLSOA_Dec_2021_PWC_for_England_and_Wales_2022_1028145039677403461.csv") %>%
  select(x, y, LSOA21CD) %>%
  semi_join(df %>% filter(is.na(LSOAXXCD)), by = "LSOA21CD")

i <- nn2(data = pts11 %>% select(-LSOA11CD),
         query = pts21 %>% select(-LSOA21CD),
         k = 1)$nn.idx[,1]

df <- left_join(df,
                data.frame(LSOA21CD = pts21$LSOA21CD, nearest = pts11$LSOA11CD[i]),
                by = "LSOA21CD") %>%
  mutate(LSOAXXCD = ifelse(!is.na(nearest), nearest, LSOAXXCD)) %>%
  select(-nearest)

```


```{r}
df |>
  filter(is.na(LSOAXXCD)) -> search

df |>
  filter(LSOA11CD %in% search$LSOA11CD |
         LSOA21CD %in% search$LSOA21CD) -> found
  
found
```


```{r}
length(unique(df$LSOAXXCD)) #34633 individual harmonised codes. 
```

```{r}
require(stringr)
df |>
  select(LSOAXXCD) |>
  filter(!duplicated(LSOAXXCD)) -> gps
data.frame(LSOAXXCD = gps$LSOAXXCD, i = 1: nrow(gps)) -> gps

df |>
  left_join(gps, by = "LSOAXXCD") |>
  mutate(i = str_pad(i, width = 7, pad = "0")) |>
  mutate(LSOAXXCD = paste0("X", i, substr(LSOAXXCD, 1, 1))) |>
  select(-i) -> df
```


```{r}
duplicated2011 <- df %>% group_by(LSOAXXCD) %>% summarise(count=n())
duplicated2011 %>% arrange(desc(count))
```
## Higher-level geographies 

Helpful to know which higher level geographies the harmonized LSOA fall within. 

```{r}
left_join(df |>
    select(LSOAXXCD, LSOA11CD, LSOA21CD,CHG21, LAD22CD, LAD22NM),
    read_csv("LSOA 2021 to BUA to LAD.csv",
             progress = FALSE,
             col_select = c(1, 3, 8, 9)) |>
    rename(LSOA21CD = 1, BUA22CD = 2, RGN22CD = 3, RGN22NM = 4), by = "LSOA21CD") -> geography
```
#Write out the data 


```{r}
geography |>
  write_csv("LSOAXXCD_full_lookup.csv")


geography |>
  select(LSOA11CD, LSOAXXCD) |>
  unique(na.rm = TRUE) |>
  arrange(LSOA11CD) |>
  write_csv("LSOA11CD_to_LSOAXXCD.csv")

geography |>
  select(LSOA21CD, LSOAXXCD) |>
  unique(na.rm = TRUE) |>
  arrange(LSOA21CD) |>
  write_csv("LSOA21CD_to_LSOAXXCD.csv")
```



## Boundary Files for the harmonised LSOA


To map the harmonised-LSOAs, create a boundary file for them, based on the 2021 Census OA boundaries, load in those boundaries and append the LSOAXXCD codes.

```{r}
require(sf)

choose.files()
map <- st_read("C://Users//sgcward//Documents//PhD//2nd paper- Risk indicator//Census Harmonised//New Updated//LSOA_(Dec_2021)_Boundaries_Generalised_Clipped_EW_(BGC).geojson") |>
  left_join(df |> select(LSOAXXCD, LSOA21CD) |> unique(), by = "LSOA21CD")  #35672 fields
```



Now aggregate the 2021 LSOA census codes to the harmonised census codes (LSOAXXCD)

```{r}
aggregate(map |> select(LSOAXXCD), by = list(map$LSOAXXCD), FUN = length) |>
  rename(LSOAXXCD = 1, nOA21 = 2) -> map

map #34633
```



Then append the higher level geographies 

```{r}
map <- map %>%
  left_join(geography, by = "LSOAXXCD")
map #35796
```



```{r}
map |>
  st_write(dsn = "LSOAXXCD_Boundaries_Generalised_Clipped_EW2.geojson")
```


```{r}
map |>
  st_write(dsn = "LSOAXXCD_Boundaries_Generalised_Clipped_EW2.shp")
```


```{r}
aggregated_map <- map %>%
  group_by(LSOAXXCD) %>%
  summarize(geometry = st_union(geometry)) %>%
  ungroup()
```

```{r}
invalid_geometries <- st_is_valid(map)

# If there are invalid geometries, attempt to make them valid
if (any(invalid_geometries)) {
  map <- st_make_valid(map)
}
```










```{r}
map |>
  st_write(dsn = "LSOAXXCD_Boundaries_Generalised_Clipped_EW2.geojson")
```

Can also save this as a shapefile for QGIS etc 

```{r}
map |>
  st_write(dsn = "LSOAXXCD_Boundaries_Generalised_Clipped_EW2.shp")
```

#########################################



```{r}
geography |>
  select(LSOAXXCD, BUA22CD,
         LAD22CD, LAD22NM,
         RGN22CD, RGN22NM, Popn) |>
  unique() -> geography
geography |>
  group_by(LSOAXXCD) |>
  summarise(Popn = sum(Popn), .groups = "keep") |>
  left_join(geography |>
    group_by(LSOAXXCD, BUA22CD) |>
    summarise(Popn = sum(Popn), .groups = "keep") |>
    group_by(LSOAXXCD) |>
    arrange(LSOAXXCD, desc(Popn)) |>
    select(-Popn) |>
    slice(1) |>
    ungroup(), by = "LSOAXXCD") |>
  left_join(geography |>
    group_by(LSOAXXCD, LAD22CD,LAD22NM) |>
    summarise(Popn = sum(Popn), .groups = "keep") |>
    group_by(LSOAXXCD) |>
    arrange(LSOAXXCD, desc(Popn)) |>
    select(-Popn) |>
    slice(1) |>
    ungroup(), by = "LSOAXXCD") |>
  left_join(geography |>
    group_by(LSOAXXCD, RGN22CD) |>
    summarise(Popn = sum(Popn), .groups = "keep") |>
    group_by(LSOAXXCD) |>
    arrange(LSOAXXCD, desc(Popn)) |>
    select(-Popn) |>
    slice(1) |>
    ungroup(), by = "LSOAXXCD") |>
  left_join(geography |>
    group_by(LSOAXXCD, RGN22NM) |>
    summarise(Popn = sum(Popn), .groups = "keep") |>
    group_by(LSOAXXCD) |>
    arrange(LSOAXXCD, desc(Popn)) |>
    select(-Popn) |>
    slice(1) |>
    ungroup(), by = "LSOAXXCD") |>
  select(LSOAXXCD, Popn, BUA22CD, LAD22CD, LAD22NM,
         RGN22CD, RGN22NM) -> geography
geography |>
  write_csv("LSOAXXCD_to_BUAXXNM_to_LADXXNM_RGNXXNM.csv")

geography
```

##################################################